Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-24 cav2017/cav2017.tex
reviewed 4.5 CEGIS with abstraction refinement
aligning notation for ADC

Linear Time Invariant (LTI) control models represent a broad class of dynamical systems
%modern embedded software, performing highly non-trivial control tasks,
with significant impact in numerous application areas as in 
the life sciences, robotics, and engineering~\cite{astrom1997computer,Franklin15}.
Control synthesis for LTI is broadly investigated, 
however the use of digital control architectures adds new challenges due to 
%exacerbated by specific artefacts,  
the effects of finite-precision arithmetic, time discretization, and quantization noise, which are typically introduced 
by Analogue-to-Digital (A/D) \todo{later ADC and DAC} and Digital-to-Analogue (D/A) conversion.  
Whilst the research on digital control is well developed~\cite{astrom1997computer}, 
automated and sound control synthesis is a challenging task,  
%Thus, the programming is a key barrier to broad adoption of digital control, 
%and requires considerable expertise.
%Whilst there is a vast literature on validation of control systems,
%the subject of control synthesis has been somewhat restricted in the properties it explores.  
%Traditionally, this focuses mainly on stability %, working typically on continuous domains
particularly when the synthesis objective goes beyond classical stability or tracking requirements. 
%\cite{DBLP:journals/corr/AbateBCCDKK16,sadabadi2016static}, or using switched
%models~\cite{DBLP:conf/emsoft/RavanbakhshS16}. 
Beyond stability, 
recently reachability tools have been developed [cite SPACEx, Phaver, and others], 
however they have not been employed towards control synthesis. 
\todo{why do we cite this paper \cite{cimatti2013parameter}?what is the relationship with our work? }
% when discretising the state-space for verification and
%parameter synthesis~\cite{cimatti2013parameter}.

@@

The tool Pessoa~\cite{mazo2010pessoa} is a MatLab toolbox that synthesizes correct-by-design
embedded control software. It is based on the description of a physical system
as an equivalent finite-state machine and on the computation of reachability properties over the finite-state 
machine. 
Based on this safety specification over state values, 
Pessoa can synthesize embedded controller software for a range of properties.  
%based on a safety specification over state values, as well as specification on characteristic functions of the controller. 
However, Pessoa does not have native inclusion of the ADC/DAC quantization error needed for
synthesizing fixed-point digital controllers.

@@

We consider models of physical plants expressed as ordinary differential equations (ODEs), 
with full state information (namely where we have access to all the model variables):  
%
\begin{align}
\label{eq:ode}
\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}, 
\end{align}
where $t \in \mathbb R_0^+$, 
where $A$ and $B$ are matrices that fully specify the continuous plant, 
and with initial states set as $x(0)$. 
Eq.~\eqref{eq:ode} is soundly discretized in time (as shown in the Appendix) into %\textcolor{red}{What is subscript ``d''?}
%
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k. 
\end{align} 
where $k \in \mathbb N$, 
and with initial state $x_{0}=x(0)$. $A_d$ and $B_d$ denote the discretized matrices of $A$ and $B$, the plant dynamics. 
Later, we will deal with the issue of variable quantization, as introduced by the ADC and DAC blocks.

@@

During any given ADC conversion, the continuous signal will be sampled in the real domain and transformed into a $\mathbb{R}\langle I_{adc},F_{adc} \rangle$ value. this sampling uses a threshold which is defined by the less significant bit ($q_{adc}=2^{-F_{adc}}$) of the ADC and some non-linearities of the circuitry. The overall conversion is
$$\mathcal{F}\langle I_{adc},F_{adc} \rangle(y(t)) = y_k : y_k \in \left[y(t)-\frac{q_{adc}}{2}\ \ \ \ y(t)+\frac{q_{adc}}{2}\right]$$.
If we denote the error in the conversion as $\nu_k=y_k-y(t) : t=nk$ then we may define some bounds for it $\nu_k \in [-\frac{q_{adc}}{2}\ \ \frac{q_{adc}}{2}]$.

We will assume, for the purposes of this analysis, that the domain of the ADC is that of the digital controller (\emph{ie} the quantizer includes any digital gain added in the code).
The process of quantization in the DAC is similar except that it is going from $\mathbb{R}\langle I_{adc},F_{adc} \rangle \rightarrow \mathbb{R}\langle I_{dac},F_{dac} \rangle$. If these domains are the same ($I_{adc}=I_{dac},F_{adc}=F_{dac}$), or if the DAC resolution in higher than the ADCs, then the DAC quantization error is equal to zero.
From the above equations we can now infer that $\nu_1 \in [-\frac{q_1}{2}\ \ \frac{q_1}{2}], \nu_2 \in [-\frac{q_2}{2}\ \ \frac{q_2}{2}] : q_1=q_{adc} \wedge q_2=q_{dac}$.
Note that these bounds hold irrespective of whether the noise is correlated, hence we may use them to over-approximate the effect of the noise on the state space progression over time.
The resulting dynamics are:
\begin{align}
\label{eq:pre_quantization}
\vec{x}_{k+1} &= \vec{A}_d\vec{x}_k+\vec{B}_d(\vec{u}_k+{\vec{\nu}_2}_k) + \vec{w}_k,\\
y_k &= \vec{C}_d \vec{x}_ k + \vec{D}_d \vec{u}_ k+{\vec{\nu}_1}_k. 
\end{align}

>>>>

Linear Time Invariant (LTI) control models represent a broad class of dynamical systems
%modern embedded software, performing highly non-trivial control tasks,
with significant impact in numerous application areas as in 
the life sciences, robotics, and engineering~\cite{astrom1997computer,Franklin15}.
Control synthesis for LTI is broadly investigated, 
however the use of digital control architectures adds new challenges due to 
%exacerbated by specific artefacts,  
the effects of finite-precision arithmetic, time discretization, and quantization noise, which are typically introduced 
by Analogue-to-Digital (ADC) and Digital-to-Analogue (DAC) conversion.  
Whilst the research on digital control is well developed~\cite{astrom1997computer}, 
automated and sound control synthesis is a challenging task,  
%Thus, the programming is a key barrier to broad adoption of digital control, 
%and requires considerable expertise.
%Whilst there is a vast literature on validation of control systems,
%the subject of control synthesis has been somewhat restricted in the properties it explores.  
%Traditionally, this focuses mainly on stability %, working typically on continuous domains
particularly when the synthesis objective goes beyond classical stability or tracking requirements. 
%\cite{DBLP:journals/corr/AbateBCCDKK16,sadabadi2016static}, or using switched
%models~\cite{DBLP:conf/emsoft/RavanbakhshS16}. 
Beyond stability, 
recently reachability tools have been developed [cite SPACEx, Phaver, and others], 
however they have not been employed towards control synthesis. 
\todo{why do we cite this paper \cite{cimatti2013parameter}?what is the relationship with our work? }
% when discretising the state-space for verification and
%parameter synthesis~\cite{cimatti2013parameter}.

@@

The tool Pessoa~\cite{mazo2010pessoa} is a MatLab toolbox that synthesizes correct-by-design
embedded control software. It is based on the description of a physical system
as an equivalent finite-state machine and on the computation of reachability properties over the finite-state 
machine. 
Based on this safety specification over state values, 
Pessoa can synthesize embedded controller software for a range of properties.  
%based on a safety specification over state values, as well as specification on characteristic functions of the controller. 
However, Pessoa does not have native inclusion of the quantization error introduced by Digital-to-Analogue/Analogue-to-Digital conversion needed for
synthesizing fixed-point digital controllers.

@@

We consider models of physical plants expressed as ordinary differential equations (ODEs), 
with full state information (namely where we have access to all the model variables):  
%
\begin{align}
\label{eq:ode}
\dot{x}(t) = Ax(t)+ B u(t), \quad x \in \mathbb{R}^{n}, u \in \mathbb{R}^m, A \in \mathbb{R}^{n \times n}, B \in \mathbb{R}^{n \times m}, 
\end{align}
where $t \in \mathbb R_0^+$, 
where $A$ and $B$ are matrices that fully specify the continuous plant, 
and with initial states set as $x(0)$. 
Eq.~\eqref{eq:ode} is soundly discretized in time (as shown in the Appendix) into %\textcolor{red}{What is subscript ``d''?}
%
\begin{align}
\label{eq:plant}
x_{k+1} = A_d x_k+ B_d u_k. 
\end{align} 
where $k \in \mathbb N$, 
and with initial state $x_{0}=x(0)$. $A_d$ and $B_d$ denote the discretized matrices of $A$ and $B$, the plant dynamics. 
Later, we will deal with the issue of variable quantization, as introduced by the ADC/DAC conversion blocks.

@@

During any given ADC conversion, the continuous signal will be sampled in the real domain and transformed into a $\mathbb{R}\langle I_{c},F_{c} \rangle$ value. this sampling uses a threshold which is defined by the less significant bit ($q_{adc}=2^{-F_{adc}}$) of the ADC and some non-linearities of the circuitry. The overall conversion is
$$\mathcal{F}\langle I_{c},F_{c} \rangle(y(t)) = y_k : y_k \in \left[y(t)-\frac{q_{c}}{2}\ \ \ \ y(t)+\frac{q_{adc}}{2}\right]$$.
If we denote the error in the conversion as $\nu_k=y_k-y(t) : t=nk$ then we may define some bounds for it $\nu_k \in [-\frac{q_{c}}{2}\ \ \frac{q_{c}}{2}]$.

We will assume, for the purposes of this analysis, that the domain of the ADC is that of the digital controller (\emph{ie} the quantizer includes any digital gain added in the code).
The process of quantization in the DAC is similar except that it is going from $\mathbb{R}\langle I_{c},F_{c} \rangle \rightarrow \mathbb{R}\langle I_{dac},F_{dac} \rangle$. If these domains are the same ($I_{c}=I_{dac},F_{c}=F_{dac}$), or if the DAC resolution in higher than the ADCs, then the DAC quantization error is equal to zero.
From the above equations we can now infer that $\nu_1 \in [-\frac{q_1}{2}\ \ \frac{q_1}{2}], \nu_2 \in [-\frac{q_2}{2}\ \ \frac{q_2}{2}] : q_1=q_{c} \wedge q_2=q_{dac}$.
Note that these bounds hold irrespective of whether the noise is correlated, hence we may use them to over-approximate the effect of the noise on the state space progression over time.
The resulting dynamics are:
\begin{align}
\label{eq:pre_quantization}
\vec{x}_{k+1} &= \vec{A}_d\vec{x}_k+\vec{B}_d(\vec{u}_k+{\vec{\nu}_2}_k) + \vec{w}_k,\\
y_k &= \vec{C}_d \vec{x}_ k + \vec{D}_d \vec{u}_ k+{\vec{\nu}_1}_k. 
\end{align}
